---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "examples/README-",
  fig.height = 3.5,
  fig.width = 5,
  fig.align = "center"
)
```


# happyR

[![Build Status](https://travis-ci.org/Illumina/happyR.svg?branch=master)](https://travis-ci.org/Illumina/happyR)
[![codecov](https://codecov.io/gh/Illumina/happyR/branch/master/graph/badge.svg)](https://codecov.io/gh/Illumina/happyR)

Load [hap.py](https://github.com/Illumina/hap.py) results into an R data structure to enable simple 
plotting, comparisons and aggregation.

## Install

```{r install, eval=FALSE}
devtools::install_github("Illumina/happyR")
```

## Usage

```{r usage, message=FALSE, warning=FALSE}
# set up
library(happyR)
library(tidyverse, quietly = TRUE)


# define happyr samplesheet
extdata_dir <- system.file("extdata", package = "happyR")
samplesheet <- tibble::tribble(
~group_id, ~replicate_id, ~happy_prefix,
"PCR-Free", "NA12878-I30", "NA12878-I30_S1",
"PCR-Free", "NA12878-I33", "NA12878-I33_S1",
"Nano", "NA12878-R1", "NA12878-R1_S1",
"Nano", "NA12878-R2", "NA12878-R2_S1"
) %>% 
mutate(happy_prefix = sprintf("%s/%s", extdata_dir, happy_prefix))


# load hap.py results
hap_samplesheet <- read_samplesheet_(samplesheet)


# query and visualise performance metrics
summary <- extract_results(hap_samplesheet$results, table = "summary") %>% 
  inner_join(samplesheet, by = "happy_prefix") %>% 
  filter(Filter == "PASS")
  
ggplot(data = summary, aes(x = METRIC.Recall, y = METRIC.Precision, color = group_id, shape = Type)) +
  geom_point() +
  xlim(NA, 1) +
  ylim(NA, 1) +
  theme_minimal() + 
  scale_color_brewer(palette = "Set2") +
  scale_size(guide = "none") +
  ggtitle("Precision-Recall scatterplot")  
```

## System requirements

Originally developed for R v3.4.0. [Tests](https://travis-ci.org/Illumina/happyR) are run using the most recent available R versions (incl. devel) on Ubuntu (Trusty) and OS X (El Capitan) platforms. happyR has not been tested on Windows. Dependencies are listed in [DESCRIPTION](DESCRIPTION). 
