---
title: "Loading hap.py results"
author: "Mar Gonzalez-Porta"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r opts, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>")
```

happyR provides an easy framework to import hap.py results into R. The user can decide whether to upload results from a single sample, by providing a happy prefix path, or to bulk load multiple datasets via a samplesheet. In both cases, it is easy to extract relevant results using helper functions.

## Set up
```{r message=FALSE, warning=FALSE}
library(happyR)
library(tidyverse)
```

## Loading and querying individual hap.py results

```{r}
# define happy_prefix (the -o argument to hap.py, here: path/to/files/happy_demo)
extdata_dir <- system.file("extdata", package = "happyR")
happy_prefix <- sprintf("%s/happy_demo", extdata_dir)

# load hap.py results
hap_result <- read_happy(happy_prefix)
```

`hap_result` is now a single `happy_result` object with the following fields:

* `summary` (from summary.csv) - a `data.frame` with high-level ALL / PASS numbers
* `extended` (from extended.csv) - a `data.frame` with region / subtype stratified metrics
* `pr_curve` (from roc.*.csv.gz) - a `happy_roc` object (a list of `data.frames`) containing precision-recall over quality score

```{r}
class(hap_result)
names(hap_result)
```

We can query each item in our `happy_result` object using standard R syntax:

```{r}
hap_result$summary %>% head

hap_result$extended %>% head

names(hap_result$pr_curve)
# e.g. here pr_curve$INDEL_PASS maps to happy_demo.roc.Locations.INDEL.PASS.csv.gz
hap_result$pr_curve$INDEL_PASS %>% head
```

Or we can use the helper `pr_data` function to query `pr_curve` with advanced filters:

```{r}
del_pr <- pr_data(hap_result, var_type = "indel", filter = "PASS", subtype = "*")
del_pr %>% head
```

## Aggregating results from multiple samples

```{r}
# define happyr samplesheet
extdata_dir <- system.file("extdata", package = "happyR")
samplesheet <- readr::read_csv("group_id,replicate_id,happy_prefix
PCR-Free,NA12878-I30,NA12878-I30_S1
PCR-Free,NA12878-I33,NA12878-I33_S1
Nano,NA12878-R1,NA12878-R1_S1
Nano,NA12878-R2,NA12878-R2_S1
") %>% 
mutate(happy_prefix = sprintf("%s/%s", extdata_dir, happy_prefix))

samplesheet

# load hap.py results
hap_samplesheet <- read_samplesheet_(samplesheet)
# or directly from a samplesheet.csv
# hap_samplesheet <- read_samplesheet(samplesheet_path = "/path/to/happyr_samplesheet.csv")
```

`hap_samplesheet` is a `happy_samplesheet` object that contains the following fields:

- `samplesheet`: a data.frame with the original samplesheet
- `results`: a `happy_result_list` that contains individual `happy_result` objects
- `ids`: a vector of result ids

```{r}
class(hap_samplesheet)
names(hap_samplesheet)
```

We can query `samplesheet` and `ids` fields as usual: 

```{r}
hap_samplesheet$samplesheet %>% head
hap_samplesheet$ids %>% head
```

And access aggregate hap.py results with the `extract_results` function:

```{r}
summary <- extract_results(hap_samplesheet$results, table = "summary")
summary %>% head
# see the extract_results documentation for a list of possible values for the table argument
```
